Spatial distribution of environmental DNA in a nearshore marine habitat

### 2016-11-18 Manuscript/v1.0_PeerJ/carkeek_grid.tex
incorporate Ole's edits and some of Jameal's

%----------------------------------ABSTRACT-------------------------------------
\section*{Abstract}
In the face of increasing threats to biodiversity, the advancement of methods for surveying biological communities is a major priority for ecologists.
Recent advances in molecular biological technologies have made it possible to detect and sequence DNA from environmental samples (environmental DNA or eDNA); however, eDNA techniques have not yet seen widespread adoption as a routine method for biological surveillance primarily due to gaps in our understanding of the dynamics of eDNA in space and time.
In order to identify the effective spatial scale of this approach in a dynamic marine environment, we collected marine surface water samples from transects ranging from the intertidal zone to 4 kilometers from shore.
Using massively parallel sequencing of 16S amplicons, we identified a diverse community of metazoans and quantified their spatial patterns using a variety of statistical tools.
We find evidence for multiple, discrete eDNA communities in this habitat, and show that these communities decrease in similarity as they become further apart.
Offshore communities tend to be richer but less even than those inshore, though diversity was not spatially autocorrelated.
Taxon-specific relative abundance coincided with our expectations of spatial distribution in taxa lacking a microscopic, pelagic life-history stage, though most of the taxa detected do not meet these criteria.
Finally, we use carefully replicated laboratory procedures to show that laboratory treatments were remarkably similar in most cases, while allowing us to detect a faulty replicate, emphasizing the importance of replication to metabarcoding studies.
While there is much work to be done before eDNA techniques can be confidently deployed as a standard method for ecological monitoring, this study serves as a first analysis of diversity at the fine spatial scales relevant to marine ecologists and confirms the promise of eDNA in dynamic environments.

@@

\subsubsection*{Aggregate Measures of Diversity}
We calculated two measures of diversity (richness and 1-Simpson's Index) to ask if aggregate metrics of the eDNA community showed evidence of spatial patterning. Richness is a measure of the number of distinct types of organisms present and so ranges from 1 (only one taxon observed) to R, the number of taxa observed across all samples. 1-Simpson's is a measure of the evenness of the distribution of abundance of taxa in a sample and ranges from 0 to 1, with the value interpreted as the probability that two sequences randomly selected from the sample will belong to different taxa. Thus larger values of the index indicate more evenly divided communities. We calculated Moran's I for both diversity metrics to test for spatial autocorrelation. We also tested for a linear effect of log-transformed distance from shore on each measure of diversity to ask how diversity changes over this strong environmental gradient.

@@

\subsubsection*{Diversity in Space}
Sites offshore tend to be less diverse (richness) and more even (1-Simpson's D) than those inshore (Figure \ref{diversity_distance}). Mean OTU richness declined by 1.42 per 1000 meters from a mean of 17.6 taxa (95\%CI = 2.15) inshore to 11.9 taxa (95\%CI = 4.31) at offshore locations (p = 0.0415; Fig. figures/diversity.pdf). 1-Simpson's diversity, the probability that two reads chosen at random from a sample belong to different species, increased by .0666 per 1000 meters from 0.225 (95\%CI = 0.0558) to 0.491 (95\%CI = $\pm$ 0.112), indicating that sequence reads were less evenly distributed among taxa in offshore samples (p $\ll$ 0.05; Figure \ref{diversity_distance}). There was no evidence for spatial autocorrelation for any of the diversity metrics (Moran's I, p > 0.05; Figure \ref{diversity}).

@@

%----------------------------------DISCUSSION-----------------------------------
\section*{Discussion}
Indirect surveys of organismal presence are a key development in ecosystem monitoring in the face of increased anthropogenic pressure and dwindling resources for ecological research. Monitoring of organisms using environmental DNA is an especially promising method, given the rapid pace of advancement in technological innovation and cost efficiency in the field of DNA sequencing and quantification. For the first time in a marine environment, we document four key patterns: (1) communities far from one another tend to be less similar than those that are nearby, (2) distinct eDNA communities exist and are distributed in a non-random fashion, (3) diversity declines with distance from shore, and (4) spatial patterning of eDNA is associated with taxon-specific life history characteristics.

@@

\subsubsection*{(2) Distinct eDNA communities exist and are distributed in a non-random fashion}
 We demonstrate strong evidence for distinct community types and the non-random spatial patterning of those communities. While the spatial distributions of communities is surprising if one were concerned only with the macroscopic life stages of metazoans, it indeed does align with the broader view that even offshore pelagic communities are comprised of and influenced by nearshore organisms. While there was no distinct break in community types between onshore and offshore sites, there was some clustering of community types that may be explained by oceanographic features such as nearshore eddies generated by strong tidal exchange in a steep bathymetric setting \citep{Yang2010}. It would be useful to better understand such features during the period of sampling, by way of oceanographic monitoring devices.

>>>>

% \pagebreak
%----------------------------------ABSTRACT-------------------------------------
\section*{Abstract}
In the face of increasing threats to biodiversity, the advancement of methods for surveying biological communities is a major priority for ecologists.
Recent advances in molecular biological technologies have made it possible to detect and sequence DNA from environmental samples (environmental DNA or eDNA); however, eDNA techniques have not yet seen widespread adoption as a routine method for biological surveillance primarily due to gaps in our understanding of the dynamics of eDNA in space and time.
In order to identify the effective spatial scale of this approach in a dynamic marine environment, we collected marine surface water samples from transects ranging from the intertidal zone to 4 kilometers from shore.
Using massively parallel sequencing of 16S amplicons, we identified a diverse community of metazoans and quantified their spatial patterns using a variety of statistical tools.
We find evidence for multiple, discrete eDNA communities in this habitat, and show that these communities decrease in similarity as they become further apart.
Offshore communities tend to be richer but less even than those inshore, though diversity was not spatially autocorrelated.
Taxon-specific relative abundance coincided with our expectations of spatial distribution in taxa lacking a microscopic, pelagic life-history stage, though most of the taxa detected do not meet these criteria.
Finally, we use carefully replicated laboratory procedures to show that laboratory treatments were remarkably similar in most cases, while allowing us to detect a faulty replicate, emphasizing the importance of replication to metabarcoding studies.
While there is much work to be done before eDNA techniques can be confidently deployed as a standard method for ecological monitoring, this study serves as a first analysis of diversity at the fine spatial scales relevant to marine ecologists and confirms the promise of eDNA in dynamic environments.

@@

\subsubsection*{Aggregate Measures of Diversity}
We calculated two measures of diversity, richness and evenness, to ask if aggregate metrics of the eDNA community showed evidence of spatial patterning. Richness is a measure of the number of distinct types of organisms present and so ranges from 1 (only one taxon observed) to $S$, the number of taxa observed across all samples. To calculate the evenness of the distribution of abundance of taxa in a sample, we used the complement of the \citet{Simpson1949} index ($1 - \Sigma{p_{i}^2}$, where $p_i$ is the proportional abundance of taxon $i$). The values of this index ranges from 0 to 1, with the value interpreted as the probability that two sequences randomly selected from the sample will belong to different taxa; thus, larger values of the index indicate more evenly divided communities \citep{Magurran2004}. We calculated Moran's I for both diversity metrics to test for spatial autocorrelation. We also tested for a linear effect of log-transformed distance from shore on each measure of diversity to ask how diversity changes over this strong environmental gradient.

@@

\subsubsection*{Diversity in Space}
Sites offshore tend to be less rich and more even than those inshore (Figure \ref{diversity_distance}). Mean OTU richness declined by 1.42 per 1000 meters from a mean of 17.6 taxa (95\%CI = 2.15) inshore to 11.9 taxa (95\%CI = 4.31) at offshore locations (p = 0.0415; Figure \ref{diversity_distance}). Evenness, the probability that two reads chosen at random from a sample belong to different species, increased by .0666 per 1000 meters from 0.225 (95\%CI = 0.0558) to 0.491 (95\%CI = $\pm$ 0.112), indicating that sequence reads were less evenly distributed among taxa in offshore samples (p $\ll$ 0.05; Figure \ref{diversity_distance}). There was no evidence for spatial autocorrelation for any of the diversity metrics (Moran's I, p > 0.05; Figure \ref{diversity}).

@@

%----------------------------------DISCUSSION-----------------------------------
\section*{Discussion}
Indirect surveys of organismal presence are a key development in ecosystem monitoring in the face of increased anthropogenic pressure and dwindling resources for ecological research. Monitoring of organisms using environmental DNA is an especially promising method, given the rapid pace of advancement in technological innovation and cost efficiency in the field of DNA sequencing and quantification. For the first time in a marine environment, we document four key patterns: (1) eDNA communities far from one another tend to be less similar than those that are nearby, (2) distinct eDNA communities exist and are distributed in a non-random fashion, (3) diversity declines with distance from shore, and (4) spatial patterning of eDNA is associated with taxon-specific life history characteristics.

@@

\subsubsection*{(2) Distinct eDNA communities exist and are distributed in a non-random fashion}
 We demonstrate strong evidence for distinct community types and the non-random spatial patterning of those communities. While the spatial distributions of communities is surprising if one were concerned only with the macroscopic life stages of metazoans, it indeed does align with the broader view that even offshore pelagic communities are comprised of and influenced by nearshore organisms. This result underscores the idea that areas immediately offshore act as ecotones, a mixing zone of taxa characteristic of benthic and pelagic environments. While there was no distinct break in community types between onshore and offshore sites, there was some clustering of community types that may be explained by oceanographic features such as nearshore eddies generated by strong tidal exchange in a steep bathymetric setting \citep{Yang2010}. It would be useful to better understand such features during the period of sampling, by way of oceanographic monitoring devices.
